---
title: "Miles per Gallon analysis"
author: "Noelene Noone"
date: "18 May 2019"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1) #switch off the warning messages
```
# Overview

This study refers to the data published in Motor Trend in 1974 recording the fuel consumption for 32 of 1973-74 model cars.  It answers 2 questions, referring to 10 aspects of car design and performance:

* Is an Automatic or Manual transmission better for Mile Per Gallon fuel consumption?

* Quantify the Mile Per Gallon difference between automatic and manual transmissions.

# Executive Summary

A simple box plot shows that on average automatic cars has a much lower fuel consumption than manual cars.  A series of multi variate regression analysis tests determine what other independent factors also play a determining role in fuel consumption and how can these be quantified.

The conclusion quantifies the Miles per Gallon difference between automatic and manual transmissions, confirming that automatic cars have a lower fuel consumption than manual cars.  It also quantifies the independent (negative) impact the number of carborators in a car has on fuel consumption.
These two factors combine to form a linear predictive model for fuel consumption.

```{r message = FALSE, echo = FALSE} 
#list libraries and data set
library(datasets)
library(car) #for Anova function
library(knitr)
data(mtcars)
```

# Analysis
## Exploration
What does this dataset look like?

```{r eval=FALSE, echo = FALSE}
## Get to know the dataset
str(mtcars)
```

The dataset contains 11 variables with 32 observations, one for each model car.
Two variables, AM (Transmission) and VS (Engine shape) needs to be converted from 0/1 numeric values to containing factor values.  Reference to the mtcars dataset description gives the values:

- VS which is short for engine shape (0 = V-shaped, 1 = straight)

- AM which is short for the transmission (0 = automatic, 1 = manual)

```{r, echo = FALSE}
## Define factor variables
mtcars1 <- within(mtcars, {
   am <- factor(am, labels = c("automatic", "manual"))
   vs <- factor(vs, labels = c("V-shaped","straight"))
})
```

## Automatic / Manual versus Miles per Gallon
The box plot (see appendix) shows the average fuel consumption is much lower for automatic than manual cars.  It also shows Automatic cars have a lower variance in MPG than Manual cars.

```{r, echo = FALSE, eval=FALSE}
#Box plot comparison between Automatic / Manual versus Miles per Gallon
plot(mtcars1$mpg ~ mtcars1$am, type = "h", main = "MT Cars comparison plot", xlab = "Automatic or Manual", ylab = "Miles per Gallon")
```

## Determine the model

By performing a series of nested model selections - combining different predictive factors -  with likelyhood ratio tests it is possible to determine the optimal combination of variables predicting fuel consumption.

### First Round

Fit a model for AM combined with each of the other factors.  Find the significant factors by comparing the p-values for Transmission (AM) in each model.  Select models where the low p-value indicates high significance.

```{r, echo = FALSE, results='asis'}
#fit a model for am combined with each of the other factors
# find the significant factors by comparing the p-values (rounded to 4 significant decimals) for am in each
# select factors that give the lowest p-value - with rouding 0 giving a 95% significance
fit_am = lm(mpg~factor(am), data = mtcars1)
sum_am <- summary(fit_am)
fit_am_Fp <- pf(sum_am$fstatistic[1],sum_am$fstatistic[2],sum_am$fstatistic[3],lower.tail = FALSE)

fit_am_hp = update(fit_am,mpg~factor(am)+hp)
sum_hp <- summary(fit_am_hp)
fit_am_hp_Fp <- pf(sum_hp$fstatistic[1],sum_hp$fstatistic[2],sum_hp$fstatistic[3],lower.tail = FALSE)
fit_am_vs = update(fit_am,mpg~factor(am)+factor(vs))
sum_vs <- summary(fit_am_vs)
fit_am_vs_Fp <- pf(sum_vs$fstatistic[1],sum_vs$fstatistic[2],sum_vs$fstatistic[3],lower.tail = FALSE)
fit_am_cyl = update(fit_am,mpg~factor(am)+cyl)
sum_cyl <- summary(fit_am_cyl)
fit_am_cyl_Fp <- pf(sum_cyl$fstatistic[1],sum_cyl$fstatistic[2],sum_cyl$fstatistic[3],lower.tail = FALSE)
fit_am_disp = update(fit_am,mpg~factor(am)+disp)
sum_disp <- summary(fit_am_disp)
fit_am_disp_Fp <- pf(sum_disp$fstatistic[1],sum_disp$fstatistic[2],sum_disp$fstatistic[3],lower.tail = FALSE)
fit_am_drat = update(fit_am,mpg~factor(am)+drat)
sum_drat <- summary(fit_am_drat)
fit_am_drat_Fp <- pf(sum_drat$fstatistic[1],sum_drat$fstatistic[2],sum_drat$fstatistic[3],lower.tail = FALSE)
fit_am_wt = update(fit_am,mpg~factor(am)+wt)
sum_wt <- summary(fit_am_wt)
fit_am_wt_Fp <- pf(sum_wt$fstatistic[1],sum_wt$fstatistic[2],sum_wt$fstatistic[3],lower.tail = FALSE)
fit_am_qsec = update(fit_am,mpg~factor(am)+qsec)
sum_qsec <- summary(fit_am_qsec)
fit_am_qsec_Fp <- pf(sum_qsec$fstatistic[1],sum_qsec$fstatistic[2],sum_qsec$fstatistic[3],lower.tail = FALSE)
fit_am_gear = update(fit_am,mpg~factor(am)+gear)
sum_gear <- summary(fit_am_gear)
fit_am_gear_Fp <- pf(sum_gear$fstatistic[1],sum_gear$fstatistic[2],sum_gear$fstatistic[3],lower.tail = FALSE)
fit_am_carb = update(fit_am,mpg~factor(am)+carb)
sum_carb <- summary(fit_am_carb)
fit_am_carb_Fp <- pf(sum_carb$fstatistic[1],sum_carb$fstatistic[2],sum_carb$fstatistic[3],lower.tail = FALSE)

two_stats <- rbind(
  c("AM and HP",round(summary(fit_am_hp)$coef[2,4],digits=3),round(summary(fit_am_hp)$coef[3,4],digits=3),round(fit_am_hp_Fp,digits=3)),
  c("AM and VS",round(summary(fit_am_vs)$coef[2,4],digits=3),round(summary(fit_am_vs)$coef[3,4],digits=3),round(fit_am_vs_Fp,digits=3)),
  c("AM and CYL",round(summary(fit_am_cyl)$coef[2,4],digits=3),round(summary(fit_am_cyl)$coef[3,4],digits=3),round(fit_am_cyl_Fp,digits=3)),
  c("AM and DISP",round(summary(fit_am_disp)$coef[2,4],digits=3),round(summary(fit_am_disp)$coef[3,4],digits=3),round(fit_am_disp_Fp,digits=3)),
  c("AM and DRAT",round(summary(fit_am_drat)$coef[2,4],digits=3),round(summary(fit_am_drat)$coef[3,4],digits=3),round(fit_am_drat_Fp,digits=3)),
  c("AM and WT",round(summary(fit_am_wt)$coef[2,4],digits=3),round(summary(fit_am_wt)$coef[3,4],digits=3),round(fit_am_wt_Fp,digits=3)),
  c("AM and QSEC",round(summary(fit_am_qsec)$coef[2,4],digits=3),round(summary(fit_am_qsec)$coef[3,4],digits=3),round(fit_am_qsec_Fp,digits=3)),
  c("AM and GEAR",round(summary(fit_am_gear)$coef[2,4],digits=3),round(summary(fit_am_gear)$coef[3,4],digits=3),round(fit_am_gear_Fp,digits=3)),
  c("AM and CARB",round(summary(fit_am_carb)$coef[2,4],digits=3),round(summary(fit_am_carb)$coef[3,4],digits=3),round(fit_am_carb_Fp,digits=3)))
colnames(two_stats) <- c("Fit","p for AM","p for 2nd","p for Model")

#Display the p-values for a two factor model"
kable(two_stats, padding=2)
```

When combining AM with HP, VS, QSEC and CARB respectively, the p-value for AM is below 0.01 inidicating significance.

### Second Round

Combine two of HP, VS, QSEC and CARB each with AM. Test the significance of the factors.

```{r, echo = FALSE, results='asis'}
#Second Round
#Combine HP, VS, QSEC and CARB each with AM.

fit_am_hp_vs = update(fit_am,mpg~factor(am)+hp+factor(vs))
sum_hp_vs <- summary(fit_am_hp_vs)
fit_am_hp_vs_Fp <- pf(sum_hp_vs$fstatistic[1],sum_hp_vs$fstatistic[2],sum_hp_vs$fstatistic[3],lower.tail = FALSE)

fit_am_hp_qsec = update(fit_am,mpg~factor(am)+hp+qsec)
sum_hp_qsec <- summary(fit_am_hp_qsec)
fit_am_hp_qsec_Fp <- pf(sum_hp_qsec$fstatistic[1],sum_hp_qsec$fstatistic[2],sum_hp_qsec$fstatistic[3],lower.tail = FALSE)

fit_am_hp_carb = update(fit_am,mpg~factor(am)+hp+carb)
sum_hp_carb <- summary(fit_am_hp_carb)
fit_am_hp_carb_Fp <- pf(sum_hp_carb$fstatistic[1],sum_hp_carb$fstatistic[2],sum_hp_carb$fstatistic[3],lower.tail = FALSE)

fit_am_vs_qsec = update(fit_am,mpg~factor(am)+factor(vs)+qsec)
sum_vs_qsec <- summary(fit_am_vs_qsec)
fit_am_vs_qsec_Fp <- pf(sum_vs_qsec$fstatistic[1],sum_vs_qsec$fstatistic[2],sum_vs_qsec$fstatistic[3],lower.tail = FALSE)

fit_am_vs_carb = update(fit_am,mpg~factor(am)+factor(vs)+carb)
sum_vs_carb <- summary(fit_am_vs_carb)
fit_am_vs_carb_Fp <- pf(sum_vs_carb$fstatistic[1],sum_vs_carb$fstatistic[2],sum_vs_carb$fstatistic[3],lower.tail = FALSE)

fit_am_qsec_carb = update(fit_am,mpg~factor(am)+qsec+carb)
sum_qsec_carb <- summary(fit_am_qsec_carb)
fit_am_qsec_carb_Fp <- pf(sum_qsec_carb$fstatistic[1],sum_qsec_carb$fstatistic[2],sum_qsec_carb$fstatistic[3],lower.tail = FALSE)

tripple_stats <- rbind(
  c("AM, HP and VS",round(summary(fit_am_hp_vs)$coef[2,4],digits=3),round(summary(fit_am_hp_vs)$coef[3,4],digits=3),round(summary(fit_am_hp_vs)$coef[4,4],digits=3),round(fit_am_hp_vs_Fp,digits=3)),
  c("AM, HP and QSEC",round(summary(fit_am_hp_qsec)$coef[2,4],digits=3),round(summary(fit_am_hp_qsec)$coef[3,4],digits=3),round(summary(fit_am_hp_qsec)$coef[4,4],digits=3),round(fit_am_hp_qsec_Fp,digits=3)),
  c("AM, HP and CARB",round(summary(fit_am_hp_carb)$coef[2,4],digits=3),round(summary(fit_am_hp_carb)$coef[3,4],digits=3),round(summary(fit_am_hp_carb)$coef[4,4],digits=3),round(fit_am_hp_carb_Fp,digits=3)),
  c("AM, VS and QSEC",round(summary(fit_am_vs_qsec)$coef[2,4],digits=3),round(summary(fit_am_vs_qsec)$coef[3,4],digits=3),round(summary(fit_am_vs_qsec)$coef[4,4],digits=3),round(fit_am_vs_qsec_Fp,digits=3)),
  c("AM, VS and CARB",round(summary(fit_am_vs_carb)$coef[2,4],digits=3),round(summary(fit_am_vs_carb)$coef[3,4],digits=3),round(summary(fit_am_vs_carb)$coef[4,4],digits=3),round(fit_am_vs_carb_Fp,digits=3)),
  c("AM, QSEC and CARB",round(summary(fit_am_qsec_carb)$coef[2,4],digits=3),round(summary(fit_am_qsec_carb)$coef[3,4],digits=3),round(summary(fit_am_qsec_carb)$coef[4,4],digits=3),round(fit_am_qsec_carb_Fp,digits=3)))

colnames(tripple_stats) <- c("Fit","p for AM","p for 2nd","p for 3rd","p for Model")

#Display the p-values for a two factor model"
kable(tripple_stats, padding=2)
```

Only the AM, VS and CARB factor combination gives a the p-value below 0.01 for all three factors.  Combine this model with the four selected two factor models.

### Determine best fit linear model
Referring to the five models, use two tests to determine the best fit model:

**1. ANOVA function analysis of the variance table to test the nested models**

```{r, echo = FALSE, results='asis'}
#ANOVA analysis of the variance table to test the nested models
an1 <- anova(fit_am,fit_am_carb)
an2<-anova(fit_am,fit_am_qsec)
an3<-anova(fit_am,fit_am_hp)
an4<-anova(fit_am,fit_am_vs,fit_am_vs_carb)

ANOVA_stats <- rbind(
  c("AM and CARB",round(an1[2,6],digits=3),"Accept",".",".","."),
  c("AM and QSEC",round(an2[2,6],digits=3),"Accept",".",".","."),
  c("AM and HP",round(an3[2,6],digits=3),"Accept",".",".","."),
  c("AM and VS",round(an4[2,6],digits=3),"Accept","AM, VS and CARB",round(an4[3,6],digits=3),"Accept"))

colnames(ANOVA_stats) <- c("Model","Variance","Result","Model","Variance","Result")

#Display the Variance inflation adding a factor to AM
kable(ANOVA_stats, padding=2)
```

The Anova test did not discoutn any of the five models, all a variance below 0.01.

Follow-up: check the variance inflation when adding the second or third factor to determine the best fit.

**2. VIF function to check Variance inflation when adding a factor**

```{r, echo = FALSE, results='asis'}
#Check Variance inflation with the VIF function

#Variance inflation for fit am and carb
vif_am_carb <- vif(fit_am_carb)
#Variance inflation for fit am, and qsec
vif_am_qsec <- vif(fit_am_qsec)
#Variance inflation for fit am and hp
vif_am_hp <- vif(fit_am_hp)
#Variance inflation for fit am, and vs
vif_am_vs <- vif(fit_am_vs)
#Variance inflation for fit am, vs and carb
vif_am_vs_carb <- vif(fit_am_vs_carb)

VIF_stats <- rbind(
  c("AM",round(vif_am_carb[1],digits=3),"CARB",round(vif_am_carb[2],digits=3),".","."),
  c("AM",round(vif_am_qsec[1],digits=3),"QSEC",round(vif_am_qsec[2],digits=3),".","."),
  c("AM",round(vif_am_hp[1],digits=3),"HP",round(vif_am_hp[2],digits=3),".","."),
  c("AM",round(vif_am_vs[1],digits=3),"VS",round(vif_am_vs[2],digits=3),".","."),
  c("AM",round(vif_am_vs_carb[1],digits=3),"VS",round(vif_am_vs_carb[2],digits=3),"CARB",round(vif_am_vs_carb[3],digits=3)))

colnames(VIF_stats) <- c("Factor","Inflation","Factor","Inflation","Factor","Inflation")

#Display the Variance inflation adding a factor to AM
kable(VIF_stats, padding=2)
```

The Variance inflation factor for the fit with AM and CARB is 1.003, much lower than any of the other models.  This means there is a low correlation between AM and CARB and points to the best fit model of the five models evaluated.

## Test the model

Perform two tests on the best fit model:

**1. Is an Interaction term required ?**

Fit the model with an interaction term.

```{r, echo = FALSE, results='asis'}
#Fit a model with an interaction term.
fit_am_carb_interact = update(fit_am,mpg~factor(am)*carb)
Interact_stats <- rbind(
  c(summary(fit_am_carb_interact)$coef[4,1],summary(fit_am_carb_interact)$coef[4,2],summary(fit_am_carb_interact)$coef[4,3],summary(fit_am_carb_interact)$coef[4,4]))

colnames(Interact_stats) <- c("Estimate","Std. Error","t value","Pr(>|t|")

#Display the p-values for a two factor model"
kable(Interact_stats, padding=2)
```

Interaction term model has a low t-value and a high p-value. Fail to reject Ho: Interaction term equal to 0.  Conclude not to add the interaction term.

**2. Is there a patern left in the Residual plot**

```{r, echo = FALSE}
mtcarsplus <- mtcars1
mtcarsplus$resid = resid(fit_am_carb)
residsum <- sum(resid(fit_am_carb)) # sum of residuals should be zero
```

```{r, echo = FALSE, eval=FALSE}
## Residual plot
plot(resid~mpg, data=mtcarsplus,main = "Residual plot", xlab = "Miles per Gallon", ylab = "Residuals")
abline(h=0, col="red")
```

Once the Systematic variation has been removed from the data with the model, the residual plot (see appendix) shows no pattern in the residual variation.  

In addition the sum of the residuals = `r residsum`.  This is rounded to zero within the mathematical precision provided in R.

## Chosen model

Below the coefficients for best fit to estimate Miles Per Gallon: 

```{r, echo = FALSE}
#Produce estimates using the chosen model
summary <- summary(fit_am_carb)$coef[,1]

All <- summary[1]
Manual <- summary[2]
Carb <- summary[3]

MilesPerGallon <- rbind(c(1,round(All+ 1 * Carb),round(All + Manual + 1 * Carb)),
                       c(2,round(All+ 2 * Carb),round(All + Manual + 2 * Carb)),
                       c(3,round(All+ 3 * Carb),round(All + Manual + 3 * Carb)),
                       c(4,round(All+ 4 * Carb),round(All + Manual + 4 * Carb)),
                       c(6,round(All+ 6 * Carb),round(All + Manual + 6 * Carb)),
                       c(8,round(All+ 8 * Carb),round(All + Manual + 8 * Carb)))
colnames(MilesPerGallon) <- c("Carburetors","Automatic","Manual")  
```

MPG = **`r round(All)`** + (**`r round(Manual)`** * AM) + (**`r round(Carb)`** * CARB) where AM=0 for automatic and AM=1 for manual and CARB is the number of carburetors.

# Conclusion

The fuel consumption in Miles per Gallon can be estimated with the model diffrentiating between automatic and manual cars and taking the number of carburetors into account:

```{r, echo = FALSE, results='asis'}
#Display the p-values for a two factor model"
kable(MilesPerGallon, padding=2)
```

\newpage


## Appendix: Data and Plots referenced

**What does this dataset look like?**
```{r, echo = FALSE, eval= TRUE}
## Get to know the dataset
str(mtcars)
```

**How is MPG different between Manual and Automatic cars ?**
```{r, echo = FALSE, eval= TRUE}
#Box plot comparison between Automatic / Manual versus Miles per Gallon
plot(mtcars1$mpg ~ mtcars1$am, type = "h", main = "MT Cars comparison plot", xlab = "Automatic or Manual", ylab = "Miles per Gallon")
```

**What does the residual plot look like ?**
```{r, echo = FALSE, eval=TRUE}
## Residual plot
plot(resid~mpg, data=mtcarsplus,main = "Residual plot", xlab = "Miles per Gallon", ylab = "Residuals")
abline(h=0, col="red")
```

## Appendix of R-code used to generate the analysis document

```{r, echo = TRUE, eval= FALSE}
## Get to know the dataset
str(mtcars)
```

```{r, echo = TRUE, eval= FALSE}
## Define factor variables
mtcars1 <- within(mtcars, {
   am <- factor(am, labels = c("automatic", "manual"))
   vs <- factor(vs, labels = c("V-shaped","straight"))
})
```

```{r, echo = TRUE, eval= FALSE}
#Box plot comparison between Automatic / Manual versus Miles per Gallon
plot(mtcars1$mpg ~ mtcars1$am, type = "h", main = "MT Cars comparison plot", xlab = "Automatic or Manual", ylab = "Miles per Gallon")
```

```{r, echo = TRUE, eval=FALSE, results='asis'}
#fit a model for am combined with each of the other factors
# find the significant factors by comparing the p-values (rounded to 4 significant decimals) for am in each
# select factors that give the lowest p-value - with rouding 0 giving a 95% significance
fit_am = lm(mpg~factor(am), data = mtcars1)
sum_am <- summary(fit_am)
fit_am_Fp <- pf(sum_am$fstatistic[1],sum_am$fstatistic[2],sum_am$fstatistic[3],lower.tail = FALSE)

fit_am_hp = update(fit_am,mpg~factor(am)+hp)
sum_hp <- summary(fit_am_hp)
fit_am_hp_Fp <- pf(sum_hp$fstatistic[1],sum_hp$fstatistic[2],sum_hp$fstatistic[3],lower.tail = FALSE)
fit_am_vs = update(fit_am,mpg~factor(am)+factor(vs))
sum_vs <- summary(fit_am_vs)
fit_am_vs_Fp <- pf(sum_vs$fstatistic[1],sum_vs$fstatistic[2],sum_vs$fstatistic[3],lower.tail = FALSE)
fit_am_cyl = update(fit_am,mpg~factor(am)+cyl)
sum_cyl <- summary(fit_am_cyl)
fit_am_cyl_Fp <- pf(sum_cyl$fstatistic[1],sum_cyl$fstatistic[2],sum_cyl$fstatistic[3],lower.tail = FALSE)
fit_am_disp = update(fit_am,mpg~factor(am)+disp)
sum_disp <- summary(fit_am_disp)
fit_am_disp_Fp <- pf(sum_disp$fstatistic[1],sum_disp$fstatistic[2],sum_disp$fstatistic[3],lower.tail = FALSE)
fit_am_drat = update(fit_am,mpg~factor(am)+drat)
sum_drat <- summary(fit_am_drat)
fit_am_drat_Fp <- pf(sum_drat$fstatistic[1],sum_drat$fstatistic[2],sum_drat$fstatistic[3],lower.tail = FALSE)
fit_am_wt = update(fit_am,mpg~factor(am)+wt)
sum_wt <- summary(fit_am_wt)
fit_am_wt_Fp <- pf(sum_wt$fstatistic[1],sum_wt$fstatistic[2],sum_wt$fstatistic[3],lower.tail = FALSE)
fit_am_qsec = update(fit_am,mpg~factor(am)+qsec)
sum_qsec <- summary(fit_am_qsec)
fit_am_qsec_Fp <- pf(sum_qsec$fstatistic[1],sum_qsec$fstatistic[2],sum_qsec$fstatistic[3],lower.tail = FALSE)
fit_am_gear = update(fit_am,mpg~factor(am)+gear)
sum_gear <- summary(fit_am_gear)
fit_am_gear_Fp <- pf(sum_gear$fstatistic[1],sum_gear$fstatistic[2],sum_gear$fstatistic[3],lower.tail = FALSE)
fit_am_carb = update(fit_am,mpg~factor(am)+carb)
sum_carb <- summary(fit_am_carb)
fit_am_carb_Fp <- pf(sum_carb$fstatistic[1],sum_carb$fstatistic[2],sum_carb$fstatistic[3],lower.tail = FALSE)

two_stats <- rbind(
  c("AM and HP",round(summary(fit_am_hp)$coef[2,4],digits=3),round(summary(fit_am_hp)$coef[3,4],digits=3),round(fit_am_hp_Fp,digits=3)),
  c("AM and VS",round(summary(fit_am_vs)$coef[2,4],digits=3),round(summary(fit_am_vs)$coef[3,4],digits=3),round(fit_am_vs_Fp,digits=3)),
  c("AM and CYL",round(summary(fit_am_cyl)$coef[2,4],digits=3),round(summary(fit_am_cyl)$coef[3,4],digits=3),round(fit_am_cyl_Fp,digits=3)),
  c("AM and DISP",round(summary(fit_am_disp)$coef[2,4],digits=3),round(summary(fit_am_disp)$coef[3,4],digits=3),round(fit_am_disp_Fp,digits=3)),
  c("AM and DRAT",round(summary(fit_am_drat)$coef[2,4],digits=3),round(summary(fit_am_drat)$coef[3,4],digits=3),round(fit_am_drat_Fp,digits=3)),
  c("AM and WT",round(summary(fit_am_wt)$coef[2,4],digits=3),round(summary(fit_am_wt)$coef[3,4],digits=3),round(fit_am_wt_Fp,digits=3)),
  c("AM and QSEC",round(summary(fit_am_qsec)$coef[2,4],digits=3),round(summary(fit_am_qsec)$coef[3,4],digits=3),round(fit_am_qsec_Fp,digits=3)),
  c("AM and GEAR",round(summary(fit_am_gear)$coef[2,4],digits=3),round(summary(fit_am_gear)$coef[3,4],digits=3),round(fit_am_gear_Fp,digits=3)),
  c("AM and CARB",round(summary(fit_am_carb)$coef[2,4],digits=3),round(summary(fit_am_carb)$coef[3,4],digits=3),round(fit_am_carb_Fp,digits=3)))
colnames(two_stats) <- c("Fit","p for AM","p for 2nd","p for Model")
#Display the p-values for a two factor model"
kable(two_stats, padding=2)
```

```{r, echo = TRUE, eval=FALSE, results='asis'}
#Second Round
#Combine HP, VS, QSEC and CARB each with AM.
fit_am_hp_vs = update(fit_am,mpg~factor(am)+hp+factor(vs))
sum_hp_vs <- summary(fit_am_hp_vs)
fit_am_hp_vs_Fp <- pf(sum_hp_vs$fstatistic[1],sum_hp_vs$fstatistic[2],sum_hp_vs$fstatistic[3],lower.tail = FALSE)

fit_am_hp_qsec = update(fit_am,mpg~factor(am)+hp+qsec)
sum_hp_qsec <- summary(fit_am_hp_qsec)
fit_am_hp_qsec_Fp <- pf(sum_hp_qsec$fstatistic[1],sum_hp_qsec$fstatistic[2],sum_hp_qsec$fstatistic[3],lower.tail = FALSE)

fit_am_hp_carb = update(fit_am,mpg~factor(am)+hp+carb)
sum_hp_carb <- summary(fit_am_hp_carb)
fit_am_hp_carb_Fp <- pf(sum_hp_carb$fstatistic[1],sum_hp_carb$fstatistic[2],sum_hp_carb$fstatistic[3],lower.tail = FALSE)

fit_am_vs_qsec = update(fit_am,mpg~factor(am)+factor(vs)+qsec)
sum_vs_qsec <- summary(fit_am_vs_qsec)
fit_am_vs_qsec_Fp <- pf(sum_vs_qsec$fstatistic[1],sum_vs_qsec$fstatistic[2],sum_vs_qsec$fstatistic[3],lower.tail = FALSE)

fit_am_vs_carb = update(fit_am,mpg~factor(am)+factor(vs)+carb)
sum_vs_carb <- summary(fit_am_vs_carb)
fit_am_vs_carb_Fp <- pf(sum_vs_carb$fstatistic[1],sum_vs_carb$fstatistic[2],sum_vs_carb$fstatistic[3],lower.tail = FALSE)

fit_am_qsec_carb = update(fit_am,mpg~factor(am)+qsec+carb)
sum_qsec_carb <- summary(fit_am_qsec_carb)
fit_am_qsec_carb_Fp <- pf(sum_qsec_carb$fstatistic[1],sum_qsec_carb$fstatistic[2],sum_qsec_carb$fstatistic[3],lower.tail = FALSE)

tripple_stats <- rbind(
  c("AM, HP and VS",round(summary(fit_am_hp_vs)$coef[2,4],digits=3),round(summary(fit_am_hp_vs)$coef[3,4],digits=3),round(summary(fit_am_hp_vs)$coef[4,4],digits=3),round(fit_am_hp_vs_Fp,digits=3)),
  c("AM, HP and QSEC",round(summary(fit_am_hp_qsec)$coef[2,4],digits=3),round(summary(fit_am_hp_qsec)$coef[3,4],digits=3),round(summary(fit_am_hp_qsec)$coef[4,4],digits=3),round(fit_am_hp_qsec_Fp,digits=3)),
  c("AM, HP and CARB",round(summary(fit_am_hp_carb)$coef[2,4],digits=3),round(summary(fit_am_hp_carb)$coef[3,4],digits=3),round(summary(fit_am_hp_carb)$coef[4,4],digits=3),round(fit_am_hp_carb_Fp,digits=3)),
  c("AM, VS and QSEC",round(summary(fit_am_vs_qsec)$coef[2,4],digits=3),round(summary(fit_am_vs_qsec)$coef[3,4],digits=3),round(summary(fit_am_vs_qsec)$coef[4,4],digits=3),round(fit_am_vs_qsec_Fp,digits=3)),
  c("AM, VS and CARB",round(summary(fit_am_vs_carb)$coef[2,4],digits=3),round(summary(fit_am_vs_carb)$coef[3,4],digits=3),round(summary(fit_am_vs_carb)$coef[4,4],digits=3),round(fit_am_vs_carb_Fp,digits=3)),
  c("AM, QSEC and CARB",round(summary(fit_am_qsec_carb)$coef[2,4],digits=3),round(summary(fit_am_qsec_carb)$coef[3,4],digits=3),round(summary(fit_am_qsec_carb)$coef[4,4],digits=3),round(fit_am_qsec_carb_Fp,digits=3)))
colnames(tripple_stats) <- c("Fit","p for AM","p for 2nd","p for 3rd","p for Model")
#Display the p-values for a two factor model"
kable(tripple_stats, padding=2)
```

```{r, echo = TRUE, eval=FALSE, results='asis'}
#ANOVA analysis of the variance table to test the nested models
an1 <- anova(fit_am,fit_am_carb)
an2<-anova(fit_am,fit_am_qsec)
an3<-anova(fit_am,fit_am_hp)
an4<-anova(fit_am,fit_am_vs,fit_am_vs_carb)
ANOVA_stats <- rbind(
  c("AM and CARB",round(an1[2,6],digits=3),"Accept",".",".","."),
  c("AM and QSEC",round(an2[2,6],digits=3),"Accept",".",".","."),
  c("AM and HP",round(an3[2,6],digits=3),"Accept",".",".","."),
  c("AM and VS",round(an4[2,6],digits=3),"Accept","AM, VS and CARB",round(an4[3,6],digits=3),"Accept"))
colnames(ANOVA_stats) <- c("Model","Variance","Result","Model","Variance","Result")
#Display the Variance inflation adding a factor to AM
kable(ANOVA_stats, padding=2)
```

```{r, echo = TRUE, eval=FALSE, results='asis'}
#Check Variance inflation with the VIF function
#Variance inflation for fit am and carb
vif_am_carb <- vif(fit_am_carb)
#Variance inflation for fit am, and qsec
vif_am_qsec <- vif(fit_am_qsec)
#Variance inflation for fit am and hp
vif_am_hp <- vif(fit_am_hp)
#Variance inflation for fit am, and vs
vif_am_vs <- vif(fit_am_vs)
#Variance inflation for fit am, vs and carb
vif_am_vs_carb <- vif(fit_am_vs_carb)
VIF_stats <- rbind(
  c("AM",round(vif_am_carb[1],digits=3),"CARB",round(vif_am_carb[2],digits=3),".","."),
  c("AM",round(vif_am_qsec[1],digits=3),"QSEC",round(vif_am_qsec[2],digits=3),".","."),
  c("AM",round(vif_am_hp[1],digits=3),"HP",round(vif_am_hp[2],digits=3),".","."),
  c("AM",round(vif_am_vs[1],digits=3),"VS",round(vif_am_vs[2],digits=3),".","."),
  c("AM",round(vif_am_vs_carb[1],digits=3),"VS",round(vif_am_vs_carb[2],digits=3),"CARB",round(vif_am_vs_carb[3],digits=3)))
colnames(VIF_stats) <- c("Factor","Inflation","Factor","Inflation","Factor","Inflation")
#Display the Variance inflation adding a factor to AM
kable(VIF_stats, padding=2)
```

```{r, echo = TRUE, eval=FALSE, results='asis'}
#Fit a model with an interaction term.
fit_am_carb_interact = update(fit_am,mpg~factor(am)*carb)
Interact_stats <- rbind(
  c(summary(fit_am_carb_interact)$coef[4,1],summary(fit_am_carb_interact)$coef[4,2],summary(fit_am_carb_interact)$coef[4,3],summary(fit_am_carb_interact)$coef[4,4]))
colnames(Interact_stats) <- c("Estimate","Std. Error","t value","Pr(>|t|")
#Display the p-values for a two factor model"
kable(Interact_stats, padding=2)
```

```{r, echo = TRUE, eval=FALSE}
## Residual plot
mtcarsplus <- mtcars1
mtcarsplus$resid = resid(fit_am_carb)
plot(resid~mpg, data=mtcarsplus,main = "Residual plot", xlab = "Miles per Gallon", ylab = "Residuals")
abline(h=0, col="red")
residsum <- sum(resid(fit_am_carb)) # sum of residuals should be zero
```

```{r, echo = TRUE, eval=FALSE, results='asis'}
#Produce estimates using the chosen model
summary <- summary(fit_am_carb)$coef[,1]
All <- summary[1]
Manual <- summary[2]
Carb <- summary[3]
MilesPerGallon <- rbind(c(1,round(All+ 1 * Carb),round(All + Manual + 1 * Carb)),
                       c(2,round(All+ 2 * Carb),round(All + Manual + 2 * Carb)),
                       c(3,round(All+ 3 * Carb),round(All + Manual + 3 * Carb)),
                       c(4,round(All+ 4 * Carb),round(All + Manual + 4 * Carb)),
                       c(6,round(All+ 6 * Carb),round(All + Manual + 6 * Carb)),
                       c(8,round(All+ 8 * Carb),round(All + Manual + 8 * Carb)))
colnames(MilesPerGallon) <- c("Carburetors","Automatic","Manual")  

#Display the p-values for a two factor model"
kable(MilesPerGallon, padding=2)
```
